---
title: "Testing Logistic"
author: "Andy Barton"
date: "September 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```



```{r}
setwd("C:/Users/barto/Documents/675/Consulting")
mydata <- read.csv("California_2016_Model_data.csv", header = TRUE)

mydata$model_1 <- as.numeric(c(mydata$Model == "A. californica"))
mydata$model_2 <- as.numeric(c(mydata$Model == "L. lorquini"))
mydata$model_3 <- as.numeric(c(mydata$Model == "L. weidemeyerii"))
mydata$model_4 <- as.numeric(c(mydata$Model == "J. coenia"))

logit_1_output <- glm(Attack ~ model_1 + model_2 + model_3, data = mydata, family = "binomial")
kable(summary(logit_1_output)$coef, digits = 3)

logit_2_output <- glm(Attack ~ model_2 + model_3 + model_4, data = mydata, family = "binomial")
kable(summary(logit_2_output)$coef, digits = 3)

logit_3_output <- glm(Attack ~ model_1 + model_3 + model_4, data = mydata, family = "binomial")
kable(summary(logit_3_output)$coef, digits = 3)
```

###P-values for 

Difference between 1 and 2    
p = 0.529   
Difference between 1 and 3    
p = 0.083    
Difference between 1 and 4    
p = 0.039   

Difference between 2 and 3    
p = 0.231   
Difference between 2 and 4    
p = 0.116   

Difference between 3 and 4    
p = 0.683

##What do you think?
As we discussed, this model makes less sense conceptually than the poisson regression. Here we see that less of these differences are statistically significant at the $\alpha$ = 0.05 level, so there is really no reason to use the model.

